---
title: "UROP Week 6 GA vs DE vs PSO"
output: html_document
---
Problems: when trying DE and PSO on DIGLOM (large number of vars), algorithm blows up in the sense that it selects way too large groups of genes & phens. This then causes p-values to be NaN-s, which prevents the algorithm from executing. 

Bad fix: set threshold, I've done this in the fitness function: if too many genes/phentypes are selected, we randomly sample out 10 and drop the others.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(CCA)
source("C:/Users/szilv/OneDrive/Dokumentumok/Imperial/UROP/Week 7/UROP_week_7_data_gen.R")
source("C:/Users/szilv/OneDrive/Dokumentumok/Imperial/UROP/Week 1-2/UROP week 2 gaCCA/genealgTwoCromosom.R")
source("C:/Users/szilv/OneDrive/Dokumentumok/Imperial/UROP/Week 1-2/UROP week 2 gaCCA/DIGLOM_gene-based.cca3.2.R")
source("C:/Users/szilv/OneDrive/Dokumentumok/Imperial/UROP/Week 1-2/UROP week 2 gaCCA/DIGLOM_gaCCAgenMultipleGene_MultiplePhen_par.R")
#source("C:/Users/szilv/OneDrive/Dokumentumok/Imperial/UROP/Week 6/UROP_week_6_get_DIGLOM.R")
source("C:/Users/szilv/OneDrive/Dokumentumok/Imperial/UROP/Week 6/UROP_week_6_PSO.R")
source("C:/Users/szilv/OneDrive/Dokumentumok/Imperial/UROP/Week 6/PSO/R/psoptim.R")
source("C:/Users/szilv/OneDrive/Dokumentumok/Imperial/UROP/Week 6/UROP_week_6_DE.R")
source("C:/Users/szilv/OneDrive/Dokumentumok/Imperial/UROP/Week 6/JDDEoptim.R")
```


```{r}
#multivariate - all subsets regression
library(tidyverse)
library(caret)
library(leaps)
set.seed(10*8)
xy <- data_generator(N=10, K=10, L=10, x_prop=0.2, y_prop=0.3, verbose=TRUE)
x <- as.data.frame(xy$x)
y <- as.data.frame(xy$y)
models <- regsubsets(x[, 8]~., data = y, nvmax = 4, really.big=T)
summary(models)
```


```{r}

#making data frames
comp_sim <- function(N, K, L,reps=2, corr="small"){

  set.seed(19*8)
  if(corr=="small"){
    xy <- data_generator(N, K, L, 0.005, 0.005, var1=10, var2=10,var3=0.2, verbose=FALSE)
  }
  if(corr=="large"){
    xy <- data_generator(N, K, L, 0.005, 0.005, var1=1, var2=1,var3=1, verbose=FALSE)
  }
  

x <- xy$x
y <- xy$y 


pso_time <- system.time({res_pso <- psoCCA(x, y, zeroToOneRatioX = 250, zeroToOneRatioY = 250, it=100, reps=reps)})

de_time <- system.time({res_de <- deCCA(x, y, zeroToOneRatioX = 250, zeroToOneRatioY = 250, it=100, reps=reps)})

ga_time  <- system.time({res_ga <- gaCCA(x, y, zeroToOneRatioX = 250, zeroToOneRatioY = 250, it=100, reps=reps)})

univariate <- function(x, y){
    require(broom)
    list_x=c()
    list_y=c()
    
    for(i in 1:L){
      
      for (j in 1:K){
        
        if(summary(lm(y[, i] ~ x[, j]))$coefficients[8] <= 0.025/(K*L)){
          #print(cor(y[, i], x[, j]))
          #print(summary(lm(y[, i] ~ x[, j]))$coefficients[8])
         
          list_x = c(list_x, j)
          list_y = c(list_y, i)
        }
      }
    }
    list1 = list(
      list_x = unique(list_x),
      list_y = unique(list_y)
    )
    
    return(list1)
}

univ_time  <- system.time({res_univ <- univariate(x, y)})



find_best <- function(data){
  pvals <- rep(0, reps)
  for(i in 1:reps){
    
    #print(data[[i]]$p_value)
    pvals[i]=data[[i]]$p_value
    
  bestInd = which(pvals==min(pvals))
  if (length(bestInd)>1){
    bestInd_new <- bestInd[1]
    return(bestInd_new)
  }
  }
  
  return(bestInd)
}

ind_ga <- find_best(res_ga)
ind_de <- find_best(res_de)
ind_pso <- find_best(res_pso)

pvals <- c(NaN, NaN, NaN,  res_ga[[ind_ga]]$p_value, res_de[[ind_de]]$p_value, res_pso[[ind_pso]]$p_value)
times <- c(NaN, NaN, as.numeric(univ_time[1]), as.numeric(ga_time[1]), as.numeric(de_time[1]), as.numeric(pso_time[1]))

genes <- c()
genes[1] = toString(sort(xy$sign_genes, decreasing=FALSE))
genes[2] = toString(sort(xy$univ_genes, decreasing=FALSE))
genes[3] = toString(sort(res_univ$list_x, decreasing=FALSE))
genes[4] = toString(res_ga[[ind_ga]]$genes)
genes[5] = toString(res_de[[ind_de]]$genes)
genes[6] = toString(res_pso[[ind_pso]]$genes)

phens <- c()
phens[1] = toString(sort(xy$sign_phens, decreasing=FALSE))
phens[2] = toString(sort(xy$univ_phens, decreasing=FALSE))
phens[3] = toString(sort(res_univ$list_y, decreasing=FALSE))
phens[4] = toString(res_ga[[ind_ga]]$phens)
phens[5] = toString(res_de[[ind_de]]$phens)
phens[6] = toString(res_pso[[ind_pso]]$phens)

all_genes <- sort(as.numeric(unique(c( sort(xy$sign_genes, decreasing=FALSE),sort(xy$univ_genes, decreasing=FALSE),  sort(res_univ$list_x, decreasing=FALSE), res_ga[[ind_ga]]$genes, res_de[[ind_de]]$genes, res_pso[[ind_pso]]$genes))), decreasing=FALSE)

all_phens <- sort(as.numeric(unique(c(sort(xy$sign_phens, decreasing=FALSE),sort(xy$univ_phens, decreasing=FALSE), sort(res_univ$list_y, decreasing=FALSE), res_ga[[ind_ga]]$phens, res_de[[ind_de]]$phens, res_pso[[ind_pso]]$phens))), decreasing=FALSE)

correl=matcor(x[, all_genes],y[, all_phens])

img.matcor(correl,type=2)

alg <- c("Construction", "Univariate check", "Univariate", "Genetic Algorithm", "Differential Evolution", "Particle Swarm Optimization")

df <- data.frame(alg, pvals, genes, phens, times)
names(df) <- c( "Algorithm","Best p-value", "Genes", "Phenotypes", "Running Time")
png("C:/Users/szilv/OneDrive/Dokumentumok/Imperial/UROP/Week 7/500_large_few_log.png", height =500, width=1500)
p <- gridExtra::tableGrob(df)
gridExtra::grid.arrange(p)
dev.off()
return(df)
}
```

```{r}

data_500_large_few_log <- comp_sim(500, 500, 500, reps=10, corr="large")
#data_small
#ptime <- system.time({2*3})
#as.numeric(ptime[3])
data_500_large_few_log
```
```{r}
data_500_small_lot_log
png("C:/Users/szilv/OneDrive/Dokumentumok/Imperial/UROP/Week 7/test.png", height=200, width=1500)
p <- gridExtra::tableGrob(data_500_small_lot_log)
gridExtra::grid.arrange(p)

dev.off()

```

```{r}
#making data frames
comp_real <- function(x, y,reps=2){

set.seed(19*8)
  

pso_time <- system.time({res_pso <- psoCCA(x, y, zeroToOneRatioX = 60, zeroToOneRatioY = 10, it=100, reps=reps)})

de_time <- system.time({res_de <- deCCA(x, y, zeroToOneRatioX = 60, zeroToOneRatioY = 10, it=100, reps=reps)})

ga_time  <- system.time({res_ga <- gaCCA(x, y, zeroToOneRatioX = 60, zeroToOneRatioY = 10, it=100, reps=reps)})

univariate <- function(x, y){
    require(broom)
    list_x=c()
    list_y=c()
    
    for(i in 1:21){
      
      for (j in 1:120){
        
        if(summary(lm(y[, i] ~ x[, j]))$coefficients[8] <= 0.025/(120*21)){
          #print(cor(y[, i], x[, j]))
          #print(summary(lm(y[, i] ~ x[, j]))$coefficients[8])
          list_x = c(list_x, j)
          list_y = c(list_y, i)
        }
      }
    }
    list1 = list(
      list_x = unique(list_x),
      list_y = unique(list_y)
    )
    
    return(list1)
}

univ_time  <- system.time({res_univ <- univariate(x, y)})



find_best <- function(data){
  pvals <- rep(0, reps)
  for(i in 1:reps){
    
    #print(data[[i]]$p_value)
    pvals[i]=data[[i]]$p_value
    
  bestInd = which(pvals==min(pvals))
  if (length(bestInd)>1){
    bestInd_new <- bestInd[1]
    return(bestInd_new)
  }
  }
  
  return(bestInd)
}

ind_ga <- find_best(res_ga)
ind_de <- find_best(res_de)
ind_pso <- find_best(res_pso)

pvals <- c( NaN,  res_ga[[ind_ga]]$p_value, res_de[[ind_de]]$p_value, res_pso[[ind_pso]]$p_value)
times <- c( as.numeric(univ_time[1]), as.numeric(ga_time[1]), as.numeric(de_time[1]), as.numeric(pso_time[1]))

genes <- c()

genes[1] = toString(sort(res_univ$list_x, decreasing=FALSE))
genes[2] = toString(res_ga[[ind_ga]]$genes)
genes[3] = toString(res_de[[ind_de]]$genes)
genes[4] = toString(res_pso[[ind_pso]]$genes)


phens <- c()

phens[1] = toString(sort(res_univ$list_y, decreasing=FALSE))
phens[2] = toString(res_ga[[ind_ga]]$phens)
phens[3] = toString(res_de[[ind_de]]$phens)
phens[4] = toString(res_pso[[ind_pso]]$phens)

all_genes <- sort(as.numeric(unique(c(sort(res_univ$list_x, decreasing=FALSE), res_ga[[ind_ga]]$genes, res_de[[ind_de]]$genes, res_pso[[ind_pso]]$genes))), decreasing=FALSE)
all_phens <- sort(as.numeric(unique(c(sort(res_univ$list_y, decreasing=FALSE), res_ga[[ind_ga]]$phens, res_de[[ind_de]]$phens, res_pso[[ind_pso]]$phens))), decreasing=FALSE)

correl=matcor(x[, all_genes],y[, all_phens])

img.matcor(correl,type=2)


alg <- c( "Univariate", "Genetic Algorithm", "Differential Evolution", "Particle Swarm Optimization")

df <- data.frame(alg, pvals, genes, phens, times)
names(df) <- c( "Algorithm","Best p-value (log scale)", "Genes", "Lipids", "Running Time")
png("C:/Users/szilv/OneDrive/Dokumentumok/Imperial/UROP/Week 7/nutri.png", height =500, width=1000)
p <- gridExtra::tableGrob(df)
gridExtra::grid.arrange(p)
dev.off()
return(df)
}
```




```{r}
#data(nutrimouse)

data_nutri <-comp_real(x= nutrimouse$gene, y=nutrimouse$lipid)
data_nutri
```








```{r}

p_threshold <- function(genes, phens, max_genes, max_phens){
    g <- 0
    for (i in 2:max_genes){
      g = g + choose(genes, i)
    }
    p <- 0
    for (i in 2:max_phens){
      p = p+ choose(phens, i)
    }

  return(0.05/(g*p))
  }


```

```{r}
#filter
xy <- data_generator(1000, 1000, 1000, 0.004, 0.006, var1=1, var2=1,var3=1, verbose=TRUE)
x <- xy$x
y <- xy$y
psoCCA(x=x, y=y, zeroToOneRatioX = 1000, zeroToOneRatioY = 1000, it=100, reps=1)
?matcor
```

```{r}
set.seed(16*8)
xy <- data_generator(N=10, K=10, L=10, x_prop=0.1, y_prop=0.2, verbose=TRUE)
x <- xy$x
y <- xy$y
  
  univariate <- function(x, y){

    require(broom)
    list_x=c()
    list_y=c()
    
    foreach(i=1:10)%:%
      
      foreach(j=1:10)%do%{
        
        if(summary(lm(y[, i] ~ x[, j]))$coefficients[8] <= 0.025/(100)){
          #print(cor(y[, i], x[, j]))
          #print(summary(lm(y[, i] ~ x[, j]))$coefficients[8])
          list_x_1 <- c(list_x, j)
          list_y_1 <- c(list_y, i)
        
      }
    }
    list1 = list(
      list_x = unique(list_x_1),
      list_y = unique(list_y_1)
    )
    
    return(list1)
}
univariate(x, y)  
```



